---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(plyr)
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data
```{r process}
activity_data <- read.csv("activity.csv")
summed_by_day <- ddply(activity_data, .(date), function(x) sum(x$steps))
avg_by_interval <- ddply(activity_data, .(interval), 
                         function(x) mean(x$steps, na.rm = T))

weekdays <- filter(activity_data, wday(date) != 1 & wday(date) != 7)
weekends <- filter(activity_data, wday(date) == 1 | wday(date) == 7)

avg_by_interval_weekday <- ddply(weekdays, .(interval), 
                         function(x) mean(x$steps, na.rm = T))
avg_by_interval_weekday <- mutate(avg_by_interval_weekday, 
                                  weekday = "Weekdays")

avg_by_interval_weekend <- ddply(weekends, .(interval), 
                         function(x) mean(x$steps, na.rm = T))
avg_by_interval_weekend <- mutate(avg_by_interval_weekend, 
                                  weekday = "Weekends")

weekend_data <- rbind.fill(avg_by_interval_weekend,
                           avg_by_interval_weekday)
```


## What is mean total number of steps taken per day?

```{r histogram_1}
hist(summed_by_day$V1, 
     xlab = "Number of Steps Per Day", 
     main = "Number of Steps Per Day Frequency")
```

```{r median_mean}
mean_steps_day <- mean(summed_by_day$V1, na.rm = T)
median_steps_day <-  median(summed_by_day$V1, na.rm = T)
```

Mean Steps per day:   `r mean_steps_day`

Median Steps per day: `r median_steps_day`


## What is the average daily activity pattern?
```{r line_plot}
plot(x = avg_by_interval$interval, 
     y = avg_by_interval$V1, 
     type = "l", 
     ylab = "Number of Steps", 
     xlab = "Interval", 
     main = "Average Steps Per Time Interval")

max_interval_idx <- which.max(avg_by_interval$V1)
max_interval <- avg_by_interval[104, c('interval')]
```

Most Active Interval: `r max_interval`

## Imputing missing values
Use the average for the time interval to replace the missing values
```{r imputing}
#Find Missings
missings <- activity_data[is.na(activity_data$steps),]
missing_count <- nrow(missings)
merged <- merge(missings, avg_by_interval, by.x = "interval", by.y = "interval")

#Create data frame and set missing values to the averages
activity_data_imputed <- data.frame(activity_data)
activity_data_imputed[is.na(activity_data_imputed$steps),]$steps <- merged$V1

summed_by_day_imputed <- ddply(activity_data_imputed, .(date), 
                               function(x) sum(x$steps))

hist(summed_by_day_imputed$V1, 
     xlab = "Number of Steps Per Day", 
     main = "Number of Steps Per Day Frequency")
```

```{r imputed_stats}
mean_steps_day_imputed <- mean(summed_by_day_imputed$V1)
median_steps_day_imputed <-  median(summed_by_day_imputed$V1)

mean_diff = (mean_steps_day - mean_steps_day_imputed)/ mean_steps_day
median_diff = (median_steps_day - median_steps_day_imputed)/ median_steps_day
```

Number of missing rows: `r missing_count`

Mean Steps per day:   `r mean_steps_day_imputed`, 
Percentage Difference: `r mean_diff`

Median Steps per day: `r median_steps_day_imputed`, 
Percentage Difference: `r median_diff`

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays_vs_weekends}
xyplot(V1 ~ interval|weekday, 
       data = weekend_data, 
       type="l",
       xlab = "Interval",
       ylab = "Average Number of Steps",
       layout = c(1,2))
```
